---
title: "Homework 3"
author: "Alyssa Sharma"
date: "2023-09-18"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(Matrix)
```

# 5.1

The code chunk below is used to check my answers for question 5.1. There are 5 outputs from this code chunk. Each output corresponds to a part of question 5.1. Hand work is shown on the notebook paper section of this question.

```{r}
a1 <- c(1,2,3)
a2 <- c(4,6,8)

a <- as.matrix(cbind(a1,a2))

b1 <- c(1,1,2)
b2 <- c(3,4,5)

b <- as.matrix(cbind(b1,b2))

c1 <- c(3,5)
c2 <- c(8,4)
c3 <- c(1,0)

c <- as.matrix(cbind(c1,c2,c3))

#part 1
print("part 1")
a + b
#part 2
print("part 2")
a - b
#part 3
print("part 3")
a %*% c
#part 4
print("part 4")
a %*% t(b)
#part 5
print("part 5")
t(b) %*% a
```


# 5.4

The code chunk below is used to check my answers for question 5.4. There are 3 outputs from this code chunk, corresponding to part 1 $\textbf{Y'Y}$, part 2 $\textbf{X'X}$ and part 3 $\textbf{X'Y}$ respectively. Hand work is shown on the notebook paper section of this question.

```{r}
x <- c(8,4,0,-4,-8)
y <- c(7.8,9,10.2,11.0,11.7)

ymat <- as.matrix(y)
ytrans <- t(ymat)

ytransy <- ytrans %*% ymat

#Answer to part 1

print("part 1")
ytransy


xmat <- as.matrix(cbind(rep(1,5),x))


xtrans <- t(xmat)


xtransx <- xtrans %*% xmat

#Answer to part 2
print("part 2")
xtransx

xtransy <- xtrans %*% ymat

#Answer to Question 5.12
invxtransx_5.12 <- solve(xtransx)
invxtransx_5.12

#Answer to part 3
print("part 3")
xtransy

```


# 5.6

The code chunk below is used to check my answers for question 5.6. There are 3 outputs from this code chunk, corresponding to part 1 $\textbf{Y'Y}$, part 2 $\textbf{X'X}$ and part 3 $\textbf{X'Y}$ respectively. Hand work is shown on the notebook paper section of this question.

```{r}
x <- c(1,0,2,0,3,1,0,1,2,0)
y <- c(16,9,17,12,22,13,8,15,19,11)

ymat <- as.matrix(y)
xmat <- as.matrix(cbind(rep(1,10),y))

#part 1
print("part 1")

ybary <- t(ymat) %*% ymat
ybary

#part 2
print("part 2")
xbarx <- t(xmat) %*% xmat
xbarx

#part 3
print("part 3")
xbary <- t(xmat) %*% ymat
xbary
```

# 5.10

The code chunk below showcases the work and answers for question 5.10. The inverse of a is shown with the first output below, ainv. The inverse of b is shown in the second output below, binv.
```{r}
a1 <-  c(2,3)
a2 <- c(4, 1)

a <- as.matrix(cbind(a1, a2))
#a

print("Here is matrix a.")
a 

ainv <- solve(a)
 

print("Here is the inverse of matrix a.")
ainv

b1 <- c(4,6,10)
b2 <- c(3,5,1)
b3 <- c(2,10,6)

b <- as.matrix(cbind(b1,b2,b3))

print("Here is matrix b.")
b

binv <- solve(b)
print("Here is the inverse of matrix b.")
binv
```

# 5.12

$(X'X)^{-1}$ is depicted in the output of the code chunk below. 

```{r}
x <- c(8,4,0,-4,-8)
y <- c(7.8,9,10.2,11.0,11.7)

ymat <- as.matrix(y)
ytrans <- t(ymat)
ytransy <- ytrans %*% ymat


xmat <- as.matrix(cbind(rep(1,5),x))
xtrans <- t(xmat)


xtransx <- xtrans %*% xmat
xtransy <- xtrans %*% ymat

#Answer to Question 5.12
invxtransx_5.12 <- solve(xtransx)
print("X'X inverse:")
invxtransx_5.12

```

# 5.14

The solutions are $y_1 = 4.5$ and $y_2 = 1$. The work for this question is depicted in the outputs below. Using the R code provided by this class's website, I found the inverse of the coefficient matrix and multiplied it by the constand matrix to find the solution matrix.

```{r}
#Part A
mat5.14 <- as.matrix(cbind(c(4,2),c(7,3)))

print("part a, coefficient matrix")
mat5.14

mat5.14end <- as.matrix(c(25,12))

print("part a, constant matrix")
mat5.14end



invX <- solve(mat5.14)
#invX 

Y<- invX %*% mat5.14end
print("part b, matrix solutions")
Y
```

